A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-03-16 papers/paper/preprint.tex
Better Bayesian inference
edits esp base SLR priors on NOAA report

\documentclass[12pt]{article}

@@

% authors
\usepackage{authblk}
\title{Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym[]{cdf}{CDF}{cumulative distribution function}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{abstract}
    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
    However, growing recognition of nonstationarity motivates a forward-looking approach.
    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-sea level rise scenarios.
    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future sea level rise to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future sea level rise.
    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
\end{abstract}

@@

Estimates of flood hazard affect peoples' lives

@@

    \item Floods affect lots of people and property
    \item Climate change and exposure making floods worse
    \item Floodplain management: $T$ year flood
    \item Some cost-benefit analysis and lots of design standards / heuristics

@@

One example is floodproofing

@@

    \item Literature on floodproofing and building-scale vulnerability reduction measures
    \item This works great!
    \item But who should use it? We're still using standards.
          Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
          For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
          Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.

@@

Nonstationarity

@@

    \item These analyses all rely on \glspl{pdf}
    \item Growing evidence for nonstationarity in a wide variety of hydroclimate processes
    \item We often focus on climate change as a driver but sometimes local modifications matter just as much \citep{Merz:2014gf}
    \item Even where risk-based analyses are used, this problem recommendations
    \item Yet stationarity remains the dominant assumption because it is immortal \citep{Montanari:2014hl} or undead \citep{Serinaldi:2015bq}
    \item Bonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{msl} is clear.
    \item On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification

@@

It matters how you prepare

@@

    \item Equifinality problem: different models diverge even if they fit the historical record well
    \item Over- and under-design can each be costly \citep{DossGollin:2019}
    \item On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
          For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
    \item On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future. For example,\ldots\ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}

@@

A quick philosophical literature review of how people have approached and interpreted this problem

@@

    \item This motivates subjective approaches for designing for nonstationarity.
    \item There's a little bit of RDM-style work on floods \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
    \item Model choice in $\mathcal{M}$-open case
    \item Philosophy of Bayesian statistics and decision making -- explain subjective probability

@@

Case study

@@

@@

Plan d'attaque

@@

    \item Research questions
          \begin{description}
              \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
              \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
              \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
          \end{description}

@@

    \item  in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .

@@

We model a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
We evaluate this decision for each of many possible simulations of \gls{msl} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
In the remainder of this section we discuss data sources and models.

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is modeled.
    }\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol            & Uncertainty                                                                          & See more                     \\
        \midrule
        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
        House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=4in]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a single decision (yellow).
        For each combination of \gls{sow} and policy, the system model (blue) is used to calculate metrics (orange) describing the performance of the policy over the selected \gls{sow}.
    }\label{fig:xlrm}
\end{figure}

\subsection{Sea level rise}\label{sec:sea-level}

Deeply uncertain sea level rise is the dominant source of deep uncertainty in our analysis.
Decision makers and analysts often have access to simulations of physical variables like \gls{msl}, but usefully incorporating these simulations into decision analyses remains an area of active research.

To inform the house elevation decision, we develop an ensemble of simulations of future \gls{msl}.
Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic models using data published in \citet{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models paramaterize physical processes, particularly of ice sheet dynamics, in different ways.
For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, or \citet{bamber_slrise:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
To represent this uncertainty, we sample four \gls{rcp} scenarios.
This gives us sixteen time-varying probability density functions of \gls{msl}.
We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \gls{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
        (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
    }\label{fig:lsl-evolution}
\end{figure}

\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
Both physical model and \gls{rcp} scenario strongly inform the modeled probability distribution.\klaus{Do you want to add a sentence here comparing BRICK to the Kopp et al models?}

@@

Coastal flood hazard does not depend only on \gls{msl} but also on storm surges.
Neglecting hydrodynamics, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
In this subsection we describe our representation of storm surge.

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
The NOAA tides and currents dataset is available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we first remove the annual mean, then extract the  maximum water level.
We refer these maxima as the annual maximum storm surges $y'(t)$.

\Cref{fig:surge-obs-return}(a) shows the time series of historical annual maximum storm surges.
Major historical surges have been driven by two different mechanisms: tropical cyclones and nor'easters.
The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and nor'easters have caused surges above \SI{6}{ft}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (A):
        time series of historic storms.
        Red (yellow) arrows denote notable tropical cyclones (nor'easters).
        (B):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
    }\label{fig:surge-obs-return}
\end{figure}

We model future storm surge using a stationary \gls{gev} model:
\begin{equation}\label{eq:surge-model}
    y'(t) \sim \text{GEV}\left(\mu, \sigma, \xi \right),
\end{equation}
where $y'(t)$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively, of the \gls{gev} distribution.
This model is stationary because there is no time dependence; surges in different years are modeled as \gls{iid}.

>>>>

\documentclass[11pt]{article}

@@

% authors
\usepackage{authblk}
\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{abc}{ABC}{approximate Bayesian computation}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym[]{cdf}{CDF}{cumulative distribution function}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

\begin{abstract}
    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
    However, growing recognition of nonstationarity motivates a forward-looking approach.
    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-\gls{slr} scenarios.
    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future \gls{slr} to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future \gls{slr}.
    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
\end{abstract}

@@

Flood risk management relies on probabilistic estimates of flood hazard:

@@

    \item In many countries including the United States, flood risk management involves many different public and private sector actors but most rely on probabilistic estimates of flood hazard.
    \item Local governments can build infrastructure to reduce flood risk; they often prioritize these by considering a reduction in damages from a design storm (\eg, the 100 year storm) \citep{hcfcd_prioritization:2019}
    \item There is a role for industry organizations that set standards that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by FEMA. At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood). \citep{kousky_voucher:2014}
    \item Beyond flood insurance and disaster recovery there has been a federal retreat from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}

@@

Future flood hazard is deeply uncertain

@@

    \item ``Along the U.S. coastline, public infrastructure and \$1 trillion in national wealth held in coastal real estate are threatened by rising sea levels, higher storm surges, and the ongoing increase in high tide flooding'' \citep{reidmiller_reportinbrief:2018}
    \item Nonstationarity also affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}
    \item For coastal flooding, \gls{fema} guidance suggests using up to date estimates of \gls{msl} \citep{fema_slr:2016}, but no projections, and stationary estimates of storm surge \citep{fema_ffa:2016}, although the analyst at each location has considerable flexibility
    \item Ditto extreme rainfall \citep{atlas14_texas:2018} or floods \citep{bulletin17c:2019}
    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}

@@

Methods for decision making under deep uncertainty emphasize exploring many possible futures

@@

    \item deep uncertainty describes ``known unknowns'' where there is not a single objectively true model \citep{walker_deep:2013,Walker:2013gi,lempert_complex:2002}
    \item This is related to equifinality where different models fit the historical record equally well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that
    \item \citet{bankes:1993} draws a distinction between exploratory and consolidative modeling
    \item Decision scaling \citep{Steinschneider:2015kk}
    \item global sensitivity analysis \citep{saltelli_sensitivity:2010,herman_salib:2017,sobol_sensitivity:2001} use feed large scenario ensembles into a black box model, then analyze outputs to assess the relative importance of each input (conditional on an assumed joint distribution over inputs)
    \item MORDM \citep{kasprzyk:2013} combines exploratory modeling (exploring performance over many scenarios) and policy search (using optimization tools)  \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}. Typically ``robustness metric'' is used \citep{herman:2015,mcphail_robustness:2019}
    \item For a more complete review see \citet{marchau:2019}
    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
    \item Crucially, these approaches all emphasize exploring performance over many scenarios, then describing this performance into some set of metrics.

@@

Some critiques

@@

    \item There are always subjective  modeling decisions, but sometimes they are opaque which makes it hard to communicate them to the end user
    \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}
    \item Technical detail: joint probability distributions over high dimensional spaces are confusing. Bayesian analysis teaches us that posterior distributions of parameters are often highly correlated. If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data!
    \item Some scenarios are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019} yet this is generally not acknowledged

@@

The question of whether probability is a useful tool for decision making under deep uncertainty depends on how one interprets probability; we argue that a subjective Bayesian interpretation of probability is useful for many DMDU problems

@@

    \item We are inspired by methods and philosophy of Bayesian model choice in the ``$\mathcal{M}$-open'' case where there is no true model to identify
    \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}. This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critiqued
    \item \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}. \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
    \item We know that choosing a single ``best'' model under-estimates uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}
    \item In the absence of data that would tell us how accurate each model is via some likelihood function, we have to use the prior.
    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}. More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}

@@

Our aim is to illustrate how methods for model selection in the $\mathcal{M}$-open case can improve decision making under deep uncertainty.
    \item We want to combine multiple models in a way that is transparent, and that can be changed with low computational cost (cheap robustness checks).
    \item We are particularly motivated by problems for which simulations of some key uncertainties (particularly climate-driven ones but also population, economic growth, etc) already exist. We assume only that there is no coupling between the system model and these simulations, so they can be treated as fully exogenous.
We illustrate our approach with a didactic case study\ldots.

@@

    \item Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
    \item For example, \citet{xian_elevation:2017} uses a cost-benefit analysis to demonstrate that tailoring recommendations to the initial elevation of a structure can reduce expected costs.
    \item Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
    \item FEMA's Hazard Mitigation Grant Program, authorized under Section 404 of the 1988 Stafford Act, which awards grants after a disaster. Under the Stafford Act, projects funded by HMGP, including buyouts, must be shown to be cost-effective; typcally, up-front costs are compared to avoided losses \citep{bendor_buyouts:2020}

@@

Our primary research question is ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
We proceed as follows.

@@

@@

    \item  in \cref{sec:exploratory} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .

@@

We model one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is represented.
    }\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol            & Uncertainty                                                                          & See more                     \\
        \midrule
        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
        House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
        First year       & $t_i$             & Deterministic: 2022                                                                  &                              \\
        Last year        & $t_f$             & Deterministic: 2091 unless otherwise noted                                           &                              \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
        For each \acrshort{sow} and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (orange).
    }\label{fig:xlrm}
\end{figure}

\subsection{\gls{slr}}\label{sec:sea-level}

To inform the house elevation decision, we develop an ensemble \gls{msl} time series.
Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, and \citet{bamber_slrise:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probability density functions of \gls{msl}.
We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \acrfull{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the choices of physical model and forcing.
        (A): 90\% confidence intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
        (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
    }\label{fig:lsl-evolution}
\end{figure}

The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
We return in \cref{sec:multiple-pdf} to the problem that these multiple models poses to decision makers.

@@

We model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
In this subsection we describe the representation of storm surge in our model.

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
The NOAA tides and currents dataset is freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
For each calendar year we first remove the annual mean, then extract the maximum water level.
We refer these maxima as the annual maximum storm surges $y'(t)$.

In the historical record, major surges have been driven by two different mechanisms: tropical cyclones and Nor'easters.
This time series of annual maxima is shown in \cref{fig:surge-obs-return}(a).
The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (A):
        time series of historic storms.
        Red (yellow) arrows denote notable tropical cyclones (Nor'easters).
        (B):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
    }\label{fig:surge-obs-return}
\end{figure}

We model future storm surge using a stationary \gls{gev} model:
\begin{equation}\label{eq:surge-model}
    y'(t) \sim \text{GEV}\qty(\mu, \sigma, \xi),
\end{equation}
where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location $\mu$, scale $\sigma$, and shape $\xi$ has the probability density function
\begin{equation*}
    f(x | \mu, \sigma, \xi)= \begin{cases}
        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
    \end{cases}
\end{equation*}
This model assumes stationarity, neglecting any potential time dependence.
